---
title: "GSS challange More Data"
author: "Kiara Butler"
date: "September 28 2023"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

$H_0$ there is no significant differences in the data.

$H_A$ There are differences in the data 

```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->


We are going to use CandP helper files to change two categorical data coupon and race. Because this is a cat- car problem will first create a bar chart then we will run fav results to show graphical and numerical results

$\Huge\Chi^2=\sumfrac{(0-E)}^2-{0}$

For interference results we are going to run a Chi-Squared test which will tell us if there any differences in the data but not where. Normally we will use a fisher test to do this but in this case we have more than one two factors in "race" so will use the XY

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results
```{r}
barchartGC(~race + cappun,data = mydata)
barchartGC(~race + cappun,data=mydata, type="percent")
```

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->



## Now with The Most data (fewer variables gives more data points)
```{r}
dd2 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

From this plot it appears the white race is more strongly in favor of capital punishment then the black and other races.



### Numerical Descriptive Results
```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```
The white race is about 70 to 30 percent in favor of capital punishment. The black race is about 50 to 50 percent. The other races are a little bit higher than 50 to 50 percent.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```


## Now with More data
```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```
The white race is about 70 to 30 percent in favor of capital punishment. The black race is about 50 to 50 percent. The other races are a little bit higher than 50 to 50 percent.


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
```


```{r}
fisherXY(mydata$race,mydata$cappun)
```

## More data version

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
#fisherXY(mydata$race,mydata$cappun)
```

the p-value for the chi-square test was  0.000000000000001027 this is too small of a number, which is why we reject the null hypothesis.


# Discussion and Conclusion
The question was if there is a difference of attitude toward capitol punishment (cappun) independent of race (race)? The answer is yes, according to the data white races think very different opinions from black and other races. The p-value was 0.000000000000001027 which is too small of a number which is why we rejected the null hypothesis. 

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
